---
title: "Running Bayestraits and simulations"
author: "Maria Izabel Cavassim"
date: "8/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dependencies 
These are the dependencies required to run these analyses

```{r}
# R packages
library(phytools)
library(geiger)
library(ape)
library(dplyr)
library(ggplot2)

# Installation of Bayestraits software
# go to:
# http://www.evolution.rdg.ac.uk/BayesTraitsV3.0.2/BayesTraitsV3.0.2.html

# Data directory:
###### ----- ####
# ~/Dropbox/PRDM9_evolution_paper/Simulations/data/

# Scripts directory:
##### ----- #####
# ~/Dropbox/PRDM9_evolution_paper/Simulations/scripts/

# Zach's script for pruning the tree (so we have both PRDM9 and the gene in the root of the tree)
source("~/Dropbox/PRDM9_evolution_paper/Simulations/scripts/pruning.R")

# Molly's script for simulation of traits (presence and absence)
source("~/Dropbox/PRDM9_evolution_paper/Simulations/scripts/sim_history_v8.R")
```

## Installing Bayestraits wrapper and running the original analysis in R
I was first running the analysis using the command line approach, but I figured that it would be
easier to have everything in one environment (R), so the pruning of the trees, simulations and runs were all within R (now I am doing only one trie at time, in the cluster I do 10 tries and take the mean value of those tries).

```{r, eval = F}
# Installing the R package to take the Bayestrait software as 
# an wrapper and run the analysis within R:
#library("devtools")
#install_github("rgriff23/btw")
library("btw")

# set your directory to the directory of the Bayestraits software
# You need to download it from here:
# http://www.evolution.rdg.ac.uk/BayesTraitsV3.0.2/BayesTraitsV3.0.2.html
setwd("/Users/PM/Desktop/BayesTraitsV3.0.2-OSX")
# I have included the executable of the software in this directory:
setwd("~/Dropbox/PRDM9_evolution_paper/Simulations")

```

# Looking at the genome assembly stats
```{r}

coverage_metadata =  read.csv("/Users/PM/Desktop/339_species/stats/stats_assemblies.csv")

# modified calls Zach
speciestree_prdm9<-read.newick(file="~/Dropbox/PRDM9_evolution_paper/Simulations/data/modified_tree.txt")

PRDM9_calls = read.csv("~/Dropbox/PRDM9_evolution_paper/Simulations/data/modified_calls.txt", header = F)
PRDM9_calls = cbind(speciestree_prdm9$tip.label, PRDM9_calls)
PRDM9_calls[which(PRDM9_calls$V1 == "TRUE"), 2] <- "1"
PRDM9_calls[which(PRDM9_calls$V1 == "FALSE"), 2] <- "0"
PRDM9_calls$Order = seq(1, nrow(PRDM9_calls))

test = merge(PRDM9_calls, coverage_metadata, by.x="speciestree_prdm9$tip.label", by.y="Species")

busco_scores = read.csv("~/Dropbox/PRDM9_evolution_paper/Simulations/data/results_255_genes_292_busco.txt")
busco_scores = subset(busco_scores, busco_scores$category == "M")
busco_scores$my_species = gsub("_protein", "", busco_scores$my_species)
test = merge(test, busco_scores, by.x="speciestree_prdm9$tip.label", by.y="my_species")

#write.csv(test, "/Users/PM/Desktop/339_species/stats/contigs_stats.txt")
setdiff(PRDM9_calls$`speciestree_prdm9$tip.label`, test$`speciestree_prdm9$tip.label`)

contigs_stat = read.csv( "/Users/PM/Desktop/339_species/stats/contigs_stats.txt", sep="\t")
contigs_stat %>%
  group_by(Order_tree, my_values) %>%
  arrange(my_values,desc(contig_N50)) -> bla

species_zeros = as.character(bla[which(bla$Order_tree == "1"), 2]$speciestree_prdm9.tip.label[1])
species_zeros = c(species_zeros, as.character(bla[which(bla$Order_tree == "2"), 2]$speciestree_prdm9.tip.label[1:3]))
#species_zeros = c(species_zeros, "Betta_splendens") 
#species_zeros = c(species_zeros, "Anabas_testudineus")
species_zeros = c(species_zeros,  as.character(bla[which(bla$Order_tree == "3"), 2]$speciestree_prdm9.tip.label[1:3]))
species_zeros = c(species_zeros,  as.character(bla[which(bla$Order_tree == "4"), 2]$speciestree_prdm9.tip.label[1:3]))
species_zeros = c(species_zeros,  as.character(bla[which(bla$Order_tree == "5"), 2]$speciestree_prdm9.tip.label[1:3]))
species_zeros = c(species_zeros,  as.character(bla[which(bla$Order_tree == "6"), 2]$speciestree_prdm9.tip.label[1:3]))
species_zeros = c(species_zeros,  as.character(bla[which(bla$Order_tree == "7"), 2]$speciestree_prdm9.tip.label[1:3]))
species_zeros = c(species_zeros,  as.character(bla[which(bla$Order_tree == "8"), 2]$speciestree_prdm9.tip.label[1:3]))
species_zeros<-species_zeros[!is.na(species_zeros)]
#species_zeros[7] <- "Nanorana_parkeri"
species_zeros_to_keep <- species_zeros

```

# Running the orginal data

Here I am only running the real data, in the next R chunk I will add the simulation part
```{r, eval = F}
##########################################
# Running our data with Bayestraits wrapper
##########################################

#read species tree:
# speciestree_prdm9<-read.newick(file="~/Dropbox/PRDM9_evolution_paper/Simulations/data/species_tree_339_changed_replacements_to_old_names.nwk")
speciestree_prdm9<-read.newick(file="~/Dropbox/PRDM9_evolution_paper/Simulations/data/modified_tree.txt")

# modified calls Zach
PRDM9_calls = read.csv("~/Dropbox/PRDM9_evolution_paper/Simulations/data/modified_calls.txt", header = F)
PRDM9_calls = cbind(speciestree_prdm9$tip.label, PRDM9_calls)
PRDM9_calls[which(PRDM9_calls$V1 == "TRUE"), 2] <- "1"
PRDM9_calls[which(PRDM9_calls$V1 == "FALSE"), 2] <- "0"

prdm9 = as.character(PRDM9_calls$V1)
names(prdm9) = PRDM9_calls$`speciestree_prdm9$tip.label`
  
# mean branch length equal to 0.1
mean_bl = 0.1/mean(speciestree_prdm9$edge.length)
median_bl = 0.1/median(speciestree_prdm9$edge.length)

########################################
###load true presence and absence matrix:
########################################
# Here we have fewer genes than 221, I need to check
presence_absence_df_filtered = read.csv("~/Dropbox/PRDM9_evolution_paper/Simulations/data/blast_results_presence_absence_genes_all_339_combined_identified_domains_metadata.csv", sep=",", row.names = 1, stringsAsFactors = FALSE)

# Excluding plytomies
presence_absence_df_filtered = subset(presence_absence_df_filtered, presence_absence_df_filtered$Species %in% speciestree_prdm9$tip.label)

# Species to exclude based on BUSCO scores
species_to_exclude = read.csv(file="~/Dropbox/PRDM9_evolution_paper/Simulations/data/species_to_exclude_busco.txt")
species_to_exclude$x = gsub("_protein", "", species_to_exclude$x)

`%notin%` <- Negate(`%in%`)
presence_absence_df_filtered = subset(presence_absence_df_filtered, presence_absence_df_filtered$Species %notin% species_to_exclude$x)

# exclude birds (except for zebra finch and chicken, called "birds" in the dataframe)
#presence_absence_df_filtered = subset(presence_absence_df_filtered, presence_absence_df_filtered$Major.clade %notin% "Aves")

# Exclude most of the "zero" species
species_ones_to_keep = presence_absence_df_filtered[presence_absence_df_filtered$Possible_assignements == "1",2]
species_to_keep = c(species_zeros_to_keep, species_ones_to_keep)

presence_absence_df_filtered = subset(presence_absence_df_filtered, presence_absence_df_filtered$Species %in% species_to_keep )

presence_absence_df_filtered = presence_absence_df_filtered[-which(presence_absence_df_filtered$Species== "Callorhinchus_milii"),]

# Exclude it from the species tree also
speciestree_prdm9 = keep.tip(speciestree_prdm9, presence_absence_df_filtered$Species)

# Checking if the tip labels differ 
setdiff(speciestree_prdm9$tip.label, row.names(presence_absence_df_filtered))
setdiff(row.names(presence_absence_df_filtered),speciestree_prdm9$tip.label)

test = merge(presence_absence_df_filtered, PRDM9_calls, by.x="Species", by.y="speciestree_prdm9$tip.label")
write.csv(test, "~/Dropbox/PRDM9_evolution_paper/Simulations/data/blast_results_presence_absence_genes_all_174_combined_identified_domains_metadata.csv")

# Emptying nodes names
speciestree_prdm9$node.label<-NULL

# mean branch length equal to 0.1
mean_bl = 0.1/mean(speciestree_prdm9$edge.length)
median_bl = 0.1/median(speciestree_prdm9$edge.length)

#prdm9 = presence_absence_df_filtered$Possible_assignements
#names(prdm9) <- presence_absence_df_filtered$Species
write.tree(speciestree_prdm9,  "~/Dropbox/PRDM9_evolution_paper/species_tree_174_3species_per_zero_calls.nw")
Bayestraits_run <- function(model="XY", tries, pres_abs=presence_absence_df_filtered, tree=speciestree_prdm9, scaling_factor=mean_bl, max_rate=100, gain_rate, prdm9_status=prdm9, pruning=F, zero_incomplete=FALSE, species_root="Petromyzon_marinus Callorhinchus_milii"){
  results_final = NULL
  not_binary = NULL
  indexing = NULL
  genes = colnames(pres_abs)[6:length(pres_abs)]
  
  species_root = toString(speciestree_prdm9$tip.label, collapse = " ")
  ##########################################
  # Models transitions (Q's)
  ##########################################
  # Species closest to the root
  species = species_root 
  #species = "Petromyzon_marinus"
  # Parameters description is found in the Bayestraits (V3) manual: 
  # http://www.evolution.rdg.ac.uk/BayesTraitsV3.0.1/Files/BayesTraitsV3.Manual.pdf
  H0_with_restrictions = c("2","1",
  #"seed 4104611364",
  # adding a tag to the species closest 
  paste("AddTag AncNode", species), 
  # here we are allowing the root to be or (1,1) : both genes presents or 1,0: at least prdm9 is present
  "Fossil Node AncNode 15", 
  "MLTol 0.0000000000000000000001",
  "mltries 100",
  paste("ScaleTrees", scaling_factor))
  
  ################################
  # Alternative Full Dependent model (XY), restrictions are also applied
  ################################
  ha1 = c("3", "1", 
  #"seed 4104611364",
  paste("AddTag AncNode", species), 
  "Fossil Node AncNode 15",
  "MLTol 0.0000000000000000000001",
  "mltries 10",
  paste("ScaleTrees", scaling_factor))
  
  ################################
  # Alternative_model_prdm9 (X), restrictions are also applied
  ################################
  ha2 = c("3", "1",
  #"seed 4104611364",
  paste("AddTag AncNode", species), 
  "Fossil Node AncNode 15",
  "MLTol 0.0000000000000000000001",
  "Restrict q31=q42",
  "mltries 100",
  paste("ScaleTrees", scaling_factor))
  
  ################################
  # Alternative_model_gene (Y),restrictions are also applied
  ################################
  ha3 = c("3", "1",
  #"seed 4104611364",
  paste("AddTag AncNode", species), 
  "Fossil Node AncNode 15",
  "MLTol 0.0000000000000000000001",
  "Restrict q21=q43",
  "mltries 20", 
  paste("ScaleTrees", scaling_factor))
  
  #genes = c("ZCWPW1", "ZCWPW2")
  for(i in genes){
    #i="ZCWPW1"
    gene2 = pres_abs[,i]
    print(i)
    names(gene2) = row.names(pres_abs)
    indexing = indexing + 1
            
    # Ordering the presence and absence based on prdm9 order!
    gene_df = as.data.frame(cbind(prdm9_status, gene2 = gene2[names(prdm9_status)]), stringsAsFactors = FALSE)
    gene_df$species = rownames(gene_df) 
    
    #gene_df["Callorhinchus_milii", 1] <- 0  
    
    # Making doubtable calls NA's
    if(zero_incomplete == TRUE){
    #gene_df["Callorhinchus_milii", 1] <- 0  
    #gene_df["Arvicanthis_niloticus", 1] <- 1 # mammal
    #gene_df["Orycteropus_afer", 1] <- 1 # mammal
    gene_df["Rhincodon_typus", 1] <- 0
    gene_df["Amblyraja_radiata", 1] <- 0
    gene_df["Acipenser_ruthenus", 1] <- 0
    gene_df["Chanos_chanos", 1] <- 0
    gene_df["Gadus_morhua", 1] <- 0
    gene_df["Salmo_trutta", 1] <- 0
    gene_df["Pangasianodon_hypophthalmus", 1] <- 0
    gene_df["Paramormyrops_kingsleyae", 1] <- 0
    gene_df["Monodelphis_domestica", 1] <- 0
    print(gene_df)
    }
    if(zero_incomplete == FALSE){
    #gene_df["Callorhinchus_milii", 1] <- "-"  
    #gene_df["Arvicanthis_niloticus", 1] <- "-"
    #gene_df["Orycteropus_afer", 1] <- "-"
    gene_df["Rhincodon_typus", 1] <- "-"
    gene_df["Amblyraja_radiata", 1] <- "-"
    gene_df["Acipenser_ruthenus", 1] <- "-"
    gene_df["Chanos_chanos", 1] <- "-"
    gene_df["Gadus_morhua", 1] <- "-"
    gene_df["Salmo_trutta", 1] <- "-"
    gene_df["Pangasianodon_hypophthalmus", 1] <- "-"
    gene_df["Paramormyrops_kingsleyae", 1] <- "-"
    gene_df["Monodelphis_domestica", 1] <- "-"
    }
    
    ### Pruning the gene so we have both 
    ### PRDM9 and the gene present in the root of the tree
            
    # Ordering the presence and absence based on the tree species order!
    gene_df = gene_df[tree$tip.label,]
    print(gene_df)
    # No pruning of the tree!!!!
    gene_df$species = tree$tip.label
    new_tree = tree
   
    # Checks of the pruning
    # Check 1: check if we excluded any species
    print("The number of tips:")
    print(length(new_tree$tip.label))
            
    # Check 2: 
    # Checking if the gene is variable
    binary = TRUE
    if(sum(as.numeric(gene_df[,2])) == nrow(gene_df)){
      print("present in all")
      binary = FALSE
      not_binary[indexing] = gene2
    }
         
    # Check 3:
    # Check the species excluded, if excluded   
    if(binary == TRUE){
      print(setdiff(tree$tip.label, new_tree$tip.label))
              
      colnames(gene_df) <- c("prdm9", "gene2", "species")
        
      # reordering the dataframe
      gene_df = gene_df[ ,c(3, 1, 2)]
        
      # emptying the node label names (otherwise the software complains)      
      new_tree$node.label<-NULL
              
      # Testing the null
      # To silent the output do silent = T
      H0 <- bayestraits(gene_df, new_tree, H0_with_restrictions, remove_files = T, silent=F)
      H0.results <- H0$Log$results
              
      #prdm9_new = as.character(gene_df$prdm9)
      #names(prdm9_new) = gene_df$species
      #gene_new = as.character(gene_df$gene2)
      #names(gene_new) = gene_df$species
              
      # Testing the alternative models
      if(model=="XY"){  
          # # Testing the alternative (XY)
          Ha <- bayestraits(gene_df, new_tree, ha1, remove_files = T, silent=F)
          Ha.results <- Ha$Log$results
          print("Alternative model results")
          print(Ha.results)
                
          # Likelihood ratio test null versus XY
          LRT.XY = 2*(Ha.results$Lh -(H0.results$Lh))
          ### Here I changed the degrees of freedom, since we are not restricting parameters any longer!!
          p.XY = 1 - pchisq(LRT.XY, df = 4) 
          print("p-value using Baystraits's method is:")
          print(p.XY)
          
          # Saving chi-square tests 
          attachs = H0.results[,c(2,3,4,5,6)] 
          print("these are the attachs")
          print(attachs)
          colnames(attachs) = c("Lik_H0", "alpha1" , "beta1", "alpha2", "beta2")
          results = cbind(Ha.results, attachs)
          results$Gene = i
          results$P.value = p.XY
          results$Trie = tries
          results_final = rbind(results_final, results) 
          }
      if(model=="X"){
          # # Testing the alternative (X)
          Ha.X <- bayestraits(gene_df, new_tree, ha2, remove_files = T, silent=F)
          Ha.X.results <- Ha.X$Log$results
                
          # Likelihood ratio test null versus X
          LRT.X = 2*(Ha.X.results$Lh -(H0.results$Lh))
          p.X = 1 - pchisq(LRT.X, df = 2)
          print(p.X)
                
          # Saving chi-square tests
          print(H0.results)
          attachs = H0.results[,c(2,3,4,5,6)] 
          print("these are the attachs")
          print(attachs)
          colnames(attachs) = c("Lik_H0", "alpha1" , "beta1", "alpha2", "beta2")
          results = cbind(Ha.X.results, attachs)
          results$Gene = i
          results$P.value = p.X
          results$Trie = tries
          results_final = rbind(results_final, results)
          }
      if(model=="Y"){
          # # Testing the alternative (X)
          Ha.Y <- bayestraits(gene_df, new_tree, ha3, remove_files = T, silent=F)
          Ha.Y.results <- Ha.Y$Log$results
                
          # Likelihood ratio test null versus Y
          LRT.Y = 2*(Ha.Y.results$Lh -(H0.results$Lh))
          p.Y = 1 - pchisq(LRT.Y, df = 2)
          print(p.Y)
          # Saving chi-square tests
          attachs = H0.results[,c(2,3,4,5,6)] 
          colnames(attachs) = c("Lik_H0", "alpha1" , "beta1", "alpha2", "beta2")
          results = cbind(Ha.Y.results, attachs)
          results$Gene = i
          results$P.value = p.Y
          results$Trie = tries
          results_final = rbind(results_final, results)
          }
        }
      }
    results_final
}

##########################################################
# All models, treating inclomplete species as missing data 
##########################################################
# Do 10 tries and take the mean
# Note that I am NOT pruning the tree any longer!!
original_results_XY_missing_2 = NULL
original_results_X_missing_2 = NULL
original_results_Y_missing_2 = NULL

for(i in 1:1){ # 1:10 if mltries == 1
  # testing all models
  trie_XY = Bayestraits_run(model="XY", tries=i, zero_incomplete = FALSE)
  #trie_X = Bayestraits_run(model="X", tries=i, zero_incomplete = FALSE)
  #trie_Y = Bayestraits_run(model="Y", tries=i, zero_incomplete = FALSE)
  
  # saving all results
  original_results_XY_missing_2 = rbind(original_results_XY_missing_2,trie_XY)
  #original_results_X_missing_2 = rbind(original_results_X_missing_2,trie_X)
  #original_results_Y_missing_2 = rbind(original_results_Y_missing_2,trie_Y)
}

original_results_XY_sig_missing = subset(original_results_XY_missing_2, original_results_XY_missing_2$P.value <= 0.05)
###
# Test for Uniformity of p-values
ks.test(original_results_XY_missing$P.value, "punif")

##### 
# Estimating FDR
# Estimating Q:
p = original_results_XY_missing$P.value

#library("devtools")
#install_github("jdstorey/qvalue")
library(qvalue)
qobj <- qvalue(p = p)

qvalues <- qobj$qvalues
pi0 <- qobj$pi0
lfdr <- qobj$lfdr

summary(qobj)
hist(qobj)
plot(qobj)

####
# Q-Q plots
# the points
expected <- -log10( (n:1)/(n+1) )
observed <- sort(-log10(p))

df = original_results_XY_missing
df = df[df$Gene != "PRDM9",]
df = df[order(df$P.value, decreasing = T),]
# df by log p-value
n=nrow(df)
df$expected = -log10((n:1)/(n+1))
df$observed = -log10(df$P.value)
df$significant = ifelse(df$P.value < 0.05, "p-value < 0.05", "Not Sig")
library("ggrepel")
p1 <- ggplot(df, aes(x=expected, y=observed)) + theme_bw() + geom_abline(intercept = 0, slope = 1) + 
  xlab(expression(paste("Expected ", -log[10](p)))) +  
  ylab(expression(paste("Observed ", -log[10](p)))) +
  geom_point(aes(color = significant)) +
  scale_color_manual(values = c("black", "red")) +
  theme_bw(base_size = 12) + theme(legend.position = "bottom") +
  geom_text_repel(
    data = subset(df, P.value < 0.05),
    aes(label = Gene),
    size = 4,
    box.padding = unit(0.2, "lines"),
    point.padding = unit(0.2, "lines")
  ) + ggtitle("Q-Q plot")


# Histogram analyses
p2 <-ggplot(df, aes(x=P.value)) + geom_histogram(color="black", fill="white", bins = 20) + theme_bw() + ggtitle("P-value distribution XY")
#library(gridExtra)
grid.arrange(p1,p2)


original_results_X_sig_missing = subset(original_results_X_missing, original_results_X_missing$P.value <= 0.05)
original_results_Y_sig_missing = subset(original_results_Y_missing, original_results_Y_missing$P.value <= 0.05)

# Saving significant files
write.table(original_results_XY_sig_missing, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_XY_sig_missing_uncertain_species.csv")
write.table(original_results_X_sig_missing, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_X_sig_missing_uncertain_species.csv")
write.table(original_results_Y_sig_missing, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_Y_sig_missing_uncertain_species.csv")

library(gridExtra)
plotting_pvalues <- function(XY, X, Y){
XY_plot <-ggplot(XY, aes(x=P.value)) + geom_histogram(color="black", fill="white", bins = 20) + theme_bw() + ggtitle("P-value distribution XY")
X_plot <-ggplot(X, aes(x=P.value)) + geom_histogram(color="black", fill="white", bins = 20) + theme_bw() + ggtitle("P-value distribution gene dependent on PRDM9")
Y_plot <-ggplot(Y, aes(x=P.value)) + geom_histogram(color="black", fill="white", bins = 20) + theme_bw() + ggtitle("P-value distribution PRDM9 dependent on gene")
grid.arrange(XY_plot, X_plot, Y_plot) 
}
plotting_pvalues(original_results_XY_missing, original_results_X_missing, original_results_Y_missing)

# Saving up the genes into a dataframe
metadata = read.csv("~/Dropbox/PRDM9_evolution_paper/Simulations/data/blast_results_presence_absence_genes_all_339_combined_identified_domains_metadata.csv", sep=",", row.names = 1, stringsAsFactors = FALSE)

metadata$Species.1  = metadata$Species
significant = metadata[,original_results_XY_sig_missing$Gene]
significant = cbind(metadata[,c(1,2,3, 148)], significant)
significant = significant[speciestree_prdm9$tip.label,]

write.csv(significant, "~/Dropbox/PRDM9_evolution_paper/original_results_XY_sig_missing.csv")

metadata$Species.1  = metadata$Species
significant_X = metadata[,original_results_X_sig_missing$Gene]
significant_X = cbind(metadata[,c(1,2,3, 148)], significant_X)
significant_X = significant_X[speciestree_prdm9$tip.label,]

write.csv(significant_X, "~/Dropbox/PRDM9_evolution_paper/original_results_X_sig_missing.csv")

significant_Y = metadata[,original_results_Y_sig_missing$Gene]
significant_Y = cbind(metadata[,c(1,2,3, 148)], significant_Y)
significant_Y = significant_Y[speciestree_prdm9$tip.label,]

write.csv(significant_Y, "~/Dropbox/PRDM9_evolution_paper/original_results_Y_sig_missing.csv")



























##########################################################
# All models, treating inclomplete species as 1 (as before)
##########################################################
# Do 10 tries and take the mean
# Note that I am NOT pruning the tree any longer!!
original_results_XY_no_Callo_in_3_species = NULL
original_results_X_no_Callo_in_3_species = NULL
original_results_Y_no_Callo_in_3_species = NULL

for(i in 1:1){ # 1:10 if mltries == 1
  # testing all models
  trie_XY = Bayestraits_run(model="XY", tries=i, zero_incomplete = "not")
  trie_X = Bayestraits_run(model="X", tries=i, zero_incomplete = "not")
  trie_Y = Bayestraits_run(model="Y", tries=i, zero_incomplete = "not")
  
  # saving all results
  original_results_XY_no_Callo_in_3_species = rbind(original_results_XY_no_Callo_in_3_species,trie_XY)
  original_results_X_no_Callo_in_3_species = rbind(original_results_X_no_Callo_in_3_species,trie_X)
  original_results_Y_no_Callo_in_3_species = rbind(original_results_Y_no_Callo_in_3_species,trie_Y)
}

original_results_XY_sig_no_Callo_in_3_species = subset(original_results_XY_no_Callo_in_3_species, original_results_XY_no_Callo_in_3_species$P.value <= 0.05)
original_results_X_sig_no_Callo_in_3_species = subset(original_results_X_no_Callo_in_3_species, original_results_X_no_Callo_in_3_species$P.value <= 0.05)
original_results_Y_sig_no_Callo_in_3_species = subset(original_results_Y_no_Callo_in_3_species, original_results_Y_no_Callo_in_3_species$P.value <= 0.05)


# Saving significant files
write.table(original_results_XY_sig_with_Callo_zero, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_XY_sig_with_Callo_0.csv")
write.table(original_results_X_sig_no_Callo_in_3_species, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_X_sig_no_Callo_in_3_species.csv")
write.table(original_results_Y_sig_with_Callo_zero, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_Y_sig_with_Callo_0.csv")


# Saving up the genes into a dataframe
metadata = read.csv("~/Dropbox/PRDM9_evolution_paper/Simulations/data/blast_results_presence_absence_genes_all_339_combined_identified_domains_metadata.csv", sep=",", row.names = 1, stringsAsFactors = FALSE)
#metadata$Possible_assignements = 
  
metadata = subset(metadata, metadata$Species %in% PRDM9_calls$`speciestree_prdm9$tip.label`)

test = merge(metadata, PRDM9_calls, by.x="Species", by.y="speciestree_prdm9$tip.label")

metadata$Possible_assignements  = test$V1
metadata$Species.1  = metadata$Species
significant = metadata[,original_results_X_sig_no_Callo_in_3_species$Gene]
significant = cbind(metadata[,c(1,2,3, 148)], significant)
significant = significant[speciestree_prdm9$tip.label,]

write.csv(significant, "~/Dropbox/PRDM9_evolution_paper/original_results_X_sig_no_Callo_in_3_species.csv")

plotting_pvalues( original_results_XY_no_Callo_in_2_species,  original_results_X_no_Callo_in_2_species,  original_results_Y_no_Callo_in_2_species)


##########################################################
# All models, treating inclomplete species as 1 (as before)
##########################################################
# Do 10 tries and take the mean
# Note that I am NOT pruning the tree any longer!!
original_results_XY_with_Callo_1 = NULL
original_results_X_with_Callo_1 = NULL
original_results_Y_with_Callo_1 = NULL

for(i in 1:1){ # 1:10 if mltries == 1
  # testing all models
  trie_XY = Bayestraits_run(model="XY", tries=i, zero_incomplete = "not")
  trie_X = Bayestraits_run(model="X", tries=i, zero_incomplete = "not")
  trie_Y = Bayestraits_run(model="Y", tries=i, zero_incomplete = "not")
  
  # saving all results
  original_results_XY_with_Callo_1 = rbind(original_results_XY_with_Callo_1,trie_XY)
  original_results_X_with_Callo_1 = rbind(original_results_X_with_Callo_1,trie_X)
  original_results_Y_with_Callo_1 = rbind(original_results_Y_with_Callo_1,trie_Y)
}

original_results_XY_sig_with_Callo_1 = subset(original_results_XY_with_Callo_1, original_results_XY_with_Callo_1$P.value <= 0.05)
original_results_X_sig_with_Callo_1 = subset(original_results_X_with_Callo_1, original_results_X_with_Callo_1$P.value <= 0.05)
original_results_Y_sig_with_Callo_1 = subset(original_results_Y_with_Callo_1, original_results_Y_with_Callo_1$P.value <= 0.05)

plotting_pvalues(original_results_XY_with_Callo_1, original_results_X_with_Callo_1, original_results_Y_with_Callo_1)

# Saving significant files
write.table(original_results_XY_sig_with_Callo_1, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_XY_sig_with_Callo_1.csv")
write.table(original_results_X_sig_with_Callo_1, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_X_sig_with_Callo_1.csv")
write.table(original_results_Y_sig_with_Callo_1, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_Y_sig_with_Callo_1.csv")


##########################################################
# All models, treating inclomplete species as 1 (as before)
##########################################################
# Do 10 tries and take the mean
# Note that I am NOT pruning the tree any longer!!
original_results_XY_no_Callo = NULL
original_results_X_no_Callo = NULL
original_results_Y_no_Callo = NULL

for(i in 1:1){ # 1:10 if mltries == 1
  # testing all models
  trie_XY = Bayestraits_run(model="XY", tries=i, zero_incomplete = "not")
  trie_X = Bayestraits_run(model="X", tries=i, zero_incomplete = "not")
  trie_Y = Bayestraits_run(model="Y", tries=i, zero_incomplete = "not")
  
  # saving all results
  original_results_XY_no_Callo = rbind(original_results_XY_no_Callo,trie_XY)
  original_results_X_no_Callo = rbind(original_results_X_no_Callo,trie_X)
  original_results_Y_no_Callo = rbind(original_results_Y_no_Callo,trie_Y)
}

original_results_XY_no_Callo_sig = subset(original_results_XY_no_Callo, original_results_XY_no_Callo$P.value <= 0.05)
original_results_X_no_Callo_sig = subset(original_results_X_no_Callo, original_results_X_no_Callo$P.value <= 0.05)
original_results_Y_no_Callo_sig = subset(original_results_Y_no_Callo, original_results_Y_no_Callo$P.value <= 0.05)

# Saving significant files
write.table(original_results_XY_no_Callo_sig, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_XY_no_Callo_sig.csv")
write.table(original_results_X_no_Callo_sig, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_X_no_Callo_sig.csv")
write.table(original_results_Y_no_Callo_sig, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_Y_no_Callo_sig.csv")


##########################################################
# All models, treating inclomplete species as 0 doing a second pruning of zero species
##########################################################
# Do 10 tries and take the mean
# Note that I am NOT pruning the tree any longer!!
original_results_XY_zero = NULL
original_results_X_zero = NULL
original_results_Y_zero = NULL

for(i in 1:1){ # 1:10 if mltries == 1
  # testing all models
  trie_XY = Bayestraits_run(model="XY", tries=i, zero_incomplete = TRUE)
  trie_X = Bayestraits_run(model="X", tries=i, zero_incomplete = TRUE)
  trie_Y = Bayestraits_run(model="Y", tries=i, zero_incomplete = TRUE)
  
  # saving all results
  original_results_XY_zero = rbind(original_results_XY_zero,trie_XY)
  original_results_X_zero = rbind(original_results_X_zero,trie_X)
  original_results_Y_zero = rbind(original_results_Y_zero,trie_Y)
}

original_results_XY_sig_zero = subset(original_results_XY_zero, original_results_XY_zero$P.value <= 0.05)
original_results_X_sig_zero = subset(original_results_X_zero, original_results_X_zero$P.value <= 0.05)
original_results_Y_sig_zero = subset(original_results_Y_zero, original_results_Y_zero$P.value <= 0.05)

plotting_pvalues(original_results_XY_zero, original_results_X_zero, original_results_Y_zero)

# Saving significant files
write.table(original_results_XY_sig_zero, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_XY_sig_zeros_uncertain_species.csv")
write.table(original_results_X_sig_zero, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_X_sig_zeros_uncertain_species.csv")
write.table(original_results_Y_sig_zero, "~/Dropbox/PRDM9_evolution_paper/Simulations/results/original_results_Y_sig_zeros_uncertain_species.csv")

```

# Validating calls

```{r}
### Functions for phylogenetic analysis
read.genbank = function(fl.name) {
  
  temp = readLines(fl.name)
  accessions = sapply(temp[grep("LOCUS",temp)], FUN = function(x) {
    t = strsplit(x," ")[[1]]
    t[t != ""][2]
  })
  aa.length = sapply(temp[grep("LOCUS",temp)], FUN = function(x) {
    t = strsplit(x," ")[[1]]
    t[t != ""][3]
  })
  definitions = sapply(grep("DEFINITION",temp), FUN = function(x) {
    t = strsplit(temp[x]," ")[[1]]
    t = t[t != ""]
    t = paste(t[2:length(t)],collapse=" ")
    incomplete = TRUE
    y = x
    while (incomplete) {
      if (length(grep("ACCESSION",temp[y+1])) == 0) {
        t2 = strsplit(temp[y+1]," ")[[1]]
        t2 = t2[t2 != ""]
        t = paste(c(t,t2),collapse=" ")
        y = y+1
      } else {
        incomplete = FALSE
      }
    }
    t
  })
  species = sapply(temp[grep("ORGANISM",temp)], FUN = function(x) {
    t = strsplit(x," ")[[1]]
    t = t[t != ""]
    paste(t[2:length(t)],collapse = " ")
  })
  end.pos = grep("//",temp)
  aa.seq = sapply(grep("ORIGIN",temp), FUN = function(x) {
    t = temp[(x+1):(end.pos[which(end.pos > x)[1]]-1)]
    paste(sapply(t,FUN = function(y) {
      t2 = strsplit(y," ")[[1]]
      t2 = t2[t2 != ""]
      paste(t2[2:length(t2)],collapse = "")
    }),collapse="")
  })
  rm(end.pos)
  gene.name = sapply(grep("/gene=",temp), FUN = function(x) {
    t = strsplit(temp[x]," ")[[1]]
    t = substring(strsplit(t[t!=""],"=")[[1]][2],2)
    substring(t,1,nchar(t)-1)
  })
  unique.ID = sapply(1:length(gene.name), FUN = function(x) {
    paste(c(strsplit(species[x]," ")[[1]],gene.name[x]),collapse="_")
  })
  
  print(length(accessions))
  print(length(definitions))
  print(length(gene.name))
  print(length(species))
  print(length(unique.ID))
  print(length(aa.length))
  print(length(aa.seq))
  result =  data.frame(accesion = accessions,
                       definition = definitions,
                       #gene = gene.name,
                       #species = species,
                       #unique.ID = unique.ID,
                       aa.length = aa.length,
                       aa.seq = aa.seq,
                       stringsAsFactors = FALSE, row.names = 1:length(accessions))
  
  # Tranform
  result
}



test = read.genbank("/Users/PM/Dropbox/PRDM9_evolution_paper/TOPGENECALLS/topgenes/RAD51AP2.gb")
write.csv(test, "/Users/PM/Dropbox/PRDM9_evolution_paper/TOPGENECALLS/topgenes/RAD51AP2_submit.txt")

test = read.genbank("/Users/PM/Dropbox/PRDM9_evolution_paper/TOPGENECALLS/topgenes/STRA8.gb")
write.csv(test, "/Users/PM/Dropbox/PRDM9_evolution_paper/TOPGENECALLS/topgenes/STRA8_submit.txt")

bla2 = read.genbank("/Users/PM/Dropbox/PRDM9_evolution_paper/TOPGENECALLS/topgenes/BRCA1.gb")
write.csv(bla2, "/Users/PM/Dropbox/PRDM9_evolution_paper/TOPGENECALLS/topgenes/BRCA1_submit.txt")

bla2 = read.genbank("/Users/PM/Dropbox/PRDM9_evolution_paper/TOPGENECALLS/topgenes/SYCE2.gb")
write.csv(bla2, "/Users/PM/Dropbox/PRDM9_evolution_paper/TOPGENECALLS/topgenes/SYCE2_submit.txt")

```